Profiling Walksheds

### 2024-10-18 data-raw/01-OSM-Network-and-Routing.Rmd
NECTAR presentation
Verified that the project is reproducible after updating renv

@@

library(disk.frame) # Larger-than-RAM Disk-Based Data Manipulation Framework
library(dplyr) # A Grammar of Data Manipulation
library(ggforce) # Accelerating 'ggplot2'
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3) # R Bindings for H3
library(purrr) # Functional Programming Tools
library(r5r) # Rapid Realistic Routing with 'R5'
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data objects from package {vaccHamilton}. Urban types is the way the City of Hamilton classifies different parts of the city as "Urban", "Suburban", and "Rural". The second object (`data_da_2016`) includes the geometry of Dissemination Areas and some demographic data for 2016:
```{r}
data("urban_types")
data("data_da_2016")
```

@@

Also, sample in urban and suburban zones to replicate the number of points for DAs. We will this as a validation dataset.
```{r}
# For reproducibility
seed <- 22331264
set.seed(seed)

@@

Save network as sf:
```{r}
save(hamilton_net, file = "hamilton_net.rda", compress = "xz")
```

@@

The origins are the coordinates of the DAs and the destinations the coordinates of the h3 geohashes:
```{r prepare-inputs}
# save origins in format expected by R5R (id, lon, lat)
origins_i_da <- data.frame(ID = origins_da$GeoUID, 
                        origins_da |>
                          st_transform(crs = 4326) |>
                          st_coordinates()) |>
  rename(lon = X, lat = Y, id = ID) |>
  dplyr::select(id, lon, lat)

@@

The DAs and the catchments match. Now save the data objects:
```{r}
save(urban_hammer_da, file = "urban_hammer_da.rda", compress = "xz")
save(ttm_walk_da, file = "ttm_walk_da.rda", compress = "xz")
save(walksheds_da, file = "walksheds_da.rda", compress = "xz")
save(ttm_walk_validation, file = "ttm_walk_validation.rda", compress = "xz")
save(walksheds_validation, file = "walksheds_validation.rda", compress = "xz")
```

@@

>>>>

**NOTE:** Notebook takes about 6 hours to run.

@@

library(disk.frame) # Larger-than-RAM Disk-Based Data Manipulation Framework
library(dplyr) # A Grammar of Data Manipulation
library(ggforce) # Accelerating 'ggplot2'
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3) # R Bindings for H3
library(here)
library(purrr) # Functional Programming Tools
library(r5r) # Rapid Realistic Routing with 'R5'
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data objects from package {vaccHamilton}. Urban types is the way the City of Hamilton classifies different parts of the city as "Urban", "Suburban", and "Rural". The second object (`data_da_2016`) includes the geometry of Dissemination Areas and some demographic data for 2016:
```{r}
data("urban_types")
data("data_da_2016")

```

@@

Also, sample in urban and suburban zones to replicate the number of points for DAs. We will this as a validation dataset. These are non-centroid origins:
```{r}
# For reproducibility
seed <- 22331264
set.seed(seed)

@@

Save network as sf:
```{r}
save(hamilton_net, 
     file = paste0(here::here(), 
                   "/data/hamilton_net.rda"), 
                   compress = "xz")
```

@@

The origins are coordinates of centroids and non-centroids points in DAs and the destinations the coordinates of the h3 geohashes:
```{r prepare-inputs}
# save origins in format expected by R5R (id, lon, lat)
origins_i_da <- data.frame(ID = origins_da$GeoUID, 
                        origins_da |>
                          st_transform(crs = 4326) |>
                          st_coordinates()) |>
  rename(lon = X, lat = Y, id = ID) |>
  dplyr::select(id, lon, lat)

@@

The DAs and the catchments match. Now save the data objects:
```{r}
save(urban_hammer_da, 
     file = paste0(here::here(), 
                   "/data/urban_hammer_da.rda"), 
     compress = "xz")
save(ttm_walk_da, 
     file = paste0(here::here(), 
                   "/data/ttm_walk_da.rda"), 
     compress = "xz")
save(walksheds_da, 
     file = paste0(here::here(), 
                   "/data/walksheds_da.rda"), 
     compress = "xz")
save(ttm_walk_validation, 
     file = paste0(here::here(), 
                   "/data/ttm_walk_validation.rda"), 
     compress = "xz")
save(walksheds_validation, 
     file = paste0(here::here(), 
                   "/data/walksheds_validation.rda"), 
     compress = "xz")
```

@@

Bind the centroid and non-centroid walksheds (da and validations, but bound could be used in more imaginative ways for cross-validation, for example):
```{r}
walksheds <- bind_rows(walksheds_da,
                       walksheds_validation) |>
  mutate(origin_type = rep(c("centroid", "non-centroid"), each = nrow(walksheds_da)))
```

### 2024-10-18 data-raw/02-OSM-Amenities.Rmd
Verified that the project is reproducible after updating renv

Load packages:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(forcats) # Tools for Working with Categorical Variables (Factors)
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3) # R Bindings for H3
library(osmdata) # Import 'OpenStreetMap' Data as Simple Features or Spatial Objects
library(purrr) # Functional Programming Tools
library(sf) # Simple Features for R
library(tidyr) # Tidy Messy Data
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

@@

Save the data objects:
```{r}
save(sustenance, file = "sustenance.rda", compress = "xz")
save(education, file = "education.rda", compress = "xz")
save(financial, file = "financial.rda", compress = "xz")
save(healthcare, file = "healthcare.rda", compress = "xz")
save(amenities_points, file = "amenities_points.rda", compress = "xz")
```

>>>>

Load packages:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(forcats) # Tools for Working with Categorical Variables (Factors)
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(h3) # R Bindings for H3
library(here) # A Simpler Way to Find Your Files
library(osmdata) # Import 'OpenStreetMap' Data as Simple Features or Spatial Objects
library(purrr) # Functional Programming Tools
library(sf) # Simple Features for R
library(tidyr) # Tidy Messy Data
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

@@

Save the data objects:
```{r}
save(sustenance, 
     file = paste0(here::here(), 
                   "/data/sustenance.rda"),
     compress = "xz")
save(education, file = paste0(here::here(), 
                   "/data/education.rda"),
     compress = "xz")
save(financial, file = paste0(here::here(), 
                   "/data/financial.rda"),
     compress = "xz")
save(healthcare, file = paste0(here::here(), 
                   "/data/healthcare.rda"),
     compress = "xz")
save(amenities_points, file = paste0(here::here(), 
                   "/data/amenities_points.rda"),
     compress = "xz")
```

### 2024-10-18 data-raw/03-Network-Attributes-by-Walkshed.Rmd
Verified that the project is reproducible after updating renv

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(igraph) # Network Analysis and Visualization
library(patchwork) # The Composer of Plots
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(sfnetworks) # Tidy Geospatial Networks
library(tidyr) # Tidy Messy Data
```

```{r}
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```

@@

Save the data object:
```{r}
save(walksheds_da_net_vars, file = "walksheds_da_net_vars.rda", compress = "xz")
```

@@

Save the data object:
```{r}
save(walksheds_validation_net_vars, file = "walksheds_validation_net_vars.rda", compress = "xz")
```

```{r}
walksheds_da_net_vars |>
  group_by(Type) |>
  skim()
```

```{r}
walksheds_validation_net_vars |>
  group_by(Type) |>
  skim()
```

@@

Save the data object:
```{r}
save(walksheds_da_motif_3_count, file = "walksheds_da_motif_3_count.rda", compress = "xz")
```

@@

```{r}
ggplot()  +
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID),
          color = "blue",
          fill = "blue") +
  geom_sf(data = urban_hammer_da |> 
            filter(GeoUID == ID), 
          fill = "red") +
  geom_sf(data = wsnet |> 
            activate(edges) |> 
            st_as_sf())
```

@@

Plot the motifs of class 10:
```{r}
ggplot() + 
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID)) +
  geom_sf(data = wsnet |>
            activate(edges) |>
            st_as_sf()) +
  geom_sf(data = motifs_class_10 |>
            activate(edges) |>
            st_as_sf(),
          color = "red") +
  geom_sf(data = motifs_class_15 |>
            activate(edges) |>
            st_as_sf(),
          color = "blue") +
  theme_void()
```

@@

Choose a walkshed with high number of motifs:
```{r}
ID <- "35250675" #

@@

@@

@@

Save object:
```{r}
save(walksheds_da_degree_dist, 
     file = "walksheds_da_degree_dist.rda", 
     compress = "xz")
```

>>>>

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggraph) # An Implementation of Grammar of Graphics for Graphs and Networks
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(igraph) # Network Analysis and Visualization
library(patchwork) # The Composer of Plots
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(sfnetworks) # Tidy Geospatial Networks
library(tidyr) # Tidy Messy Data
```

```{r}
load(file = glue::glue(here::here(), "/data/hamilton_net.rda"))
load(file = glue::glue(here::here(), "/data/urban_hammer_da.rda"))
load(file = glue::glue(here::here(), "/data/walksheds_da.rda"))
load(file = glue::glue(here::here(), "/data/walksheds_validation.rda"))
```

@@

Save the data object:
```{r}
save(walksheds_da_net_vars, 
     file = glue::glue(here::here(), "/data/walksheds_da_net_vars.rda"), 
     compress = "xz")
```

@@

Save the data object:
```{r}
save(walksheds_validation_net_vars, 
     file = glue::glue(here::here(), 
                       "/data/walksheds_validation_net_vars.rda"),
     compress = "xz")
```

Summarize. Notice that `girth` is problematic for the skim, although it can be examined with `summary()`:
```{r}
walksheds_da_net_vars |>
  select(-girth) |>
  group_by(Type) |>
  skim()
```
Summarize the walksheds for 
```{r}
walksheds_validation_net_vars |>
  select(-girth) |>
  group_by(Type) |>
  skim()
```

@@

Save the data object:
```{r}
save(walksheds_da_motif_3_count, 
     file = glue::glue(here::here(), "/data/walksheds_da_motif_3_count.rda"), compress = "xz")
```

@@

Plot chosen walkshed:
```{r}
ggplot()  +
  geom_sf(data = urban_hammer_da |> 
            filter(GeoUID == ID), 
          fill = "red") +
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID),
          color = "blue",
          fill = "blue") +
  geom_sf(data = wsnet |> 
            activate(edges) |> 
            st_as_sf())
```

@@

Plot the motifs of class 10 and 15:
```{r}
ggplot() + 
  geom_sf(data = walksheds_da |>
            filter(GeoUID == ID)) +
  geom_sf(data = wsnet |>
            activate(edges) |>
            st_as_sf()) +
  geom_sf(data = motifs_class_10 |>
            activate(edges) |>
            st_as_sf(),
          color = "red") +
  geom_sf(data = motifs_class_15 |>
            activate(edges) |>
            st_as_sf(),
          color = "blue") +
  theme_void()
```

@@

Choose a walkshed with high number of motifs:
```{r}
ID <- "35250677" #

@@

<!--
Find the motifs of class 10 and 15 in wsnet:
```{r}
iso_class_10_list <- subgraph_isomorphisms(size_3_iso[[10]], 
                                           wsnet, 
                                           method = "vf2")

@@

Almost all elements of the network are part of a motif. But the number of motifs is low, which suggests that a large number of dead ends tends to reduce the number of motifs.
-->

@@

Save object:
```{r}
save(walksheds_da_degree_dist, 
     file = glue::glue(here::here(), "/data/walksheds_da_degree_dist.rda"), 
     compress = "xz")
```

### 2024-10-18 data-raw/04-Accessibility-Calculations.Rmd
Verified that the project is reproducible after updating renv

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3)
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = "amenities_points.rda")
load(file = "urban_hammer_da.rda")
data(urban_types)
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```

@@

Save amenities by walkshed:
```{r}
save(amenities_walksheds_da, file = "amenities_walkshed_da.rda", compress = "xz")
save(amenities_walksheds_validation, file = "amenities_walkshed_validation.rda", compress = "xz")
```

@@

Distance to amenities:
```{r}
load("ttm_walk_da.rda")
```

```{r}
amenities_by_h3 <- amenities_points |>
  st_drop_geometry() |>
  group_by(h3_index) |>
  summarize(n_amenities = n(),
            .groups = "drop")
```

@@

For all destinations this is the equation:
$$
\log(Y) = 1.282 - 0.048 \cdot t
$$
For all destinations this is the equation:
$$
\log(Y) = 1.035 - 0.032 \cdot t
$$

$Y$ is the percentage of trips and $t$ is the time in minutes.

```{r}
ggplot() +
  geom_function(fun = function(x) exp(1.282)*exp(-0.048)^x,
                color = "blue") +
  geom_function(fun = function(x) exp(1.035) * exp(-0.032)^x,
                color = "red") +
  #geom_vline(xintercept =  4.37) +
  #geom_vline(xintercept = 6.02) +
  xlim(1, 15)
```

>>>>

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(h3) # R Bindings for H3
library(here) # A Simpler Way to Find Your Files
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = paste0(here::here(), 
                   "/data/amenities_points.rda"))
load(file = paste0(here::here(), 
                   "/data/urban_hammer_da.rda"))
data(urban_types)
load(file = paste0(here::here(), 
                   "/data/walksheds_da.rda"))
load(file = paste0(here::here(), 
                   "/data/walksheds_validation.rda"))
```

@@

Save amenities by walkshed:
```{r}
save(amenities_walksheds_da, 
     file = paste0(here::here(), 
                   "/data/amenities_walkshed_da.rda"),
     compress = "xz")
save(amenities_walksheds_validation, file = paste0(here::here(), 
                   "/data/amenities_walkshed_validation.rda"),
     compress = "xz")
```

@@

Distance to amenities:
```{r}
load(paste0(here::here(), 
                   "/data/ttm_walk_da.rda"))
```

Summary of amenities by h3 geometry:
```{r}
amenities_by_h3 <- amenities_points |>
  st_drop_geometry() |>
  group_by(h3_index) |>
  summarize(n_amenities = n(),
            .groups = "drop")
```

@@

For all destinations this is the equation:
$$
\log(Y) = 1.282 - 0.048 \cdot t
$$

For all destinations this is the equation:
$$
\log(Y) = 1.035 - 0.032 \cdot t
$$

$Y$ is the percentage of trips and $t$ is the time in minutes.
```{r}
ggplot() +
  geom_function(fun = function(x) exp(1.282)*exp(-0.048)^x,
                color = "blue") +
  geom_function(fun = function(x) exp(1.035) * exp(-0.032)^x,
                color = "red") +
  #geom_vline(xintercept =  4.37) +
  #geom_vline(xintercept = 6.02) +
  xlim(1, 15)
```

### 2024-10-18 data-raw/05-Walkshed-Profiles.Rmd
Verified that the project is reproducible after updating renv

Load packages:
```{r warning=FALSE, message=FALSE}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:
```{r}
load(file = "amenities_walkshed_da.rda")
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
data("urban_types")
load(file = "walksheds_da.rda")
load(file = "walksheds_da_net_vars.rda")
```

@@

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:
```{r}
d <- walksheds_da_net_vars |>
                  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)
```

We use a technique with evolutionary trees - an alternative to classification trees that is less affecte by the greediness of the algorithm. Since evolutionary trees have some randomness (for the initial values), we can try several to find whether there are patterns. In this case we try four different random seed.

Tree 1 with seed 1 (912571):
```{r}
seed1 <- sample.int(1000000, 1)
#seed1 <- 908544
set.seed(908544)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree1)
```

Tree 2 with seed 2 (94766):
```{r}
seed2 <- sample.int(1000000, 1)
#seed2 <- 669402
seed2
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree2)
```

Tree 3 with seed 3 (46125):
```{r}
seed3 <- sample.int(1000000, 1)
#seed3 <- 853251
seed3
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree3)
```

Tree 4 with seed 4 (556090)
```{r}
seed4 <- sample.int(1000000, 1)
#seed4 <- 707633
seed4
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree4)
```

Create a function to estimate the missclassification rate and the evaluation function:
```{r}
mc <- function(obj) 1 - mean(predict(obj) == d$Type)

@@

Plot the trees:
```{r}
tree1_plot <- ggparty(tree1) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

@@

Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```

@@

0.00591859212548868 & normalized_motifs_3 < 0.86734693877551 ~ "Leaf 1",
                          edge_density < 0.00591859212548868 & normalized_motifs_3 >= 0.86734693877551 ~ "Leaf 2",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 < 0.967741935483871 ~ "Leaf 3",

@@

<!--
Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```

Plot the tree with the rounded break labels:
```{r}
ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = str_replace(breaks_label, 
                                          "0.0207311636201516", 
                                          "0.021"))) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.0,
                     ids = c(1, 2, 5)) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.02,
                     ids = c(3, 4, 6, 7)) +
  geom_node_plot(gglist = list(geom_bar(aes(x = "", 
                                            fill = Type),
                                        position = position_fill()),
                               xlab("Type"),
                               theme_minimal()),
                 # draw only one label for each axis
                 shared_axis_labels = TRUE,
                 # draw line between tree and legend
                 legend_separator = FALSE
                 )
```

@@

Leaf 1:
```{r}

@@

Leaf 2 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#84451, 44785

@@

Leaf 3 (suburban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#81461, 84451

@@

Leaf 4 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#52534

@@

Add leaf labels to data:
```{r}
d <- d |>
  mutate(leaf = case_when(transitivity < 0.0207311636201516 & motifs < 378 ~ "Leaf 1",
                          transitivity < 0.0207311636201516 & motifs >= 378 ~ "Leaf 2",
                          transitivity >= 0.0207311636201516 & motifs < 1024 ~ "Leaf 3",
                          transitivity >= 0.0207311636201516 & motifs >= 1024 ~ "Leaf 4"),
         leaf = factor(leaf))
```

@@

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

To map accessibility by category of amenity we first categorize the number of amenities:
```{r}
amentities_da <-  urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE))
```

Plot:
```{r}
amentities_da |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Pivot wider:
```{r}
junk <- amentities_da |>
  pivot_wider(names_from = "Amenity_Category",
              values_from = "Number_of_Amenities")
```

>>>>

Load packages:

```{r warning=FALSE, message=FALSE}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:

```{r}
load(file = glue::glue(here::here(), "/data/amenities_walkshed_da.rda"))
load(file = glue::glue(here::here(), "/data/hamilton_net.rda"))
load(file = glue::glue(here::here(), "/data/urban_hammer_da.rda"))
data("urban_types")
load(file = glue::glue(here::here(), "/data/walksheds_da.rda"))
load(file = glue::glue(here::here(), "/data/walksheds_da_net_vars.rda"))
```

@@

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:

```{r}
d <- walksheds_da_net_vars |>
  select(-girth) |>
  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)
```

We use a technique with evolutionary trees - an alternative to classification trees that is less affected by the greediness of the algorithm. Since evolutionary trees have some randomness (for the initial values), we can try several to find whether there are patterns. In this case we try four different random seed.

Tree 1 with seed 1 (912571):

```{r}
seed1 <- sample.int(1000000, 1)
#seed1 <- 908544
set.seed(908544)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree1)
```

Tree 2 with seed 2 (94766):

```{r}
seed2 <- sample.int(1000000, 1)
#seed2 <- 669402
seed2
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree2)
```

Tree 3 with seed 3 (46125):

```{r}
seed3 <- sample.int(1000000, 1)
#seed3 <- 853251
seed3
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree3)
```

Tree 4 with seed 4 (556090)

```{r}
seed4 <- sample.int(1000000, 1)
#seed4 <- 707633
seed4
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree4)
```

Create a function to estimate the missclassification rate and the evaluation function:

```{r}
mc <- function(obj) 1 - mean(predict(obj) == d$Type)

@@

Plot the trees:

```{r}
tree1_plot <- ggparty(tree1) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

@@

Check the break labels of `tree3` to round:

```{r}
ggparty(tree3)$data$breaks_label
```

@@

0.00591859212548868 & normalized_motifs_3 \< 0.86734693877551 \~ "Leaf 1", edge_density \< 0.00591859212548868 & normalized_motifs_3 \>= 0.86734693877551 \~ "Leaf 2", edge_density \>= 0.00591859212548868 & normalized_motifs_3 \< 0.967741935483871 \~ "Leaf 3",

@@

\<!-- Check the break labels of `tree3` to round:

```{r}
ggparty(tree3)$data$breaks_label
```

Plot the tree with the rounded break labels:

```{r}
ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = str_replace(breaks_label, 
                                          "0.0207311636201516", 
                                          "0.021"))) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.0,
                     ids = c(1, 2, 5)) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.02,
                     ids = c(3, 4, 6, 7)) +
  geom_node_plot(gglist = list(geom_bar(aes(x = "", 
                                            fill = Type),
                                        position = position_fill()),
                               xlab("Type"),
                               theme_minimal()),
                 # draw only one label for each axis
                 shared_axis_labels = TRUE,
                 # draw line between tree and legend
                 legend_separator = FALSE
                 )
```

@@

Leaf 1:

```{r}

@@

Leaf 2 (urban):

```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#84451, 44785

@@

Leaf 3 (suburban):

```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#81461, 84451

@@

Leaf 4 (urban):

```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#52534

@@

Add leaf labels to data:

```{r}
d <- d |>
  mutate(leaf = case_when(transitivity < 0.0207311636201516 & motifs < 378 ~ "Leaf 1",
                          transitivity < 0.0207311636201516 & motifs >= 378 ~ "Leaf 2",
                          transitivity >= 0.0207311636201516 & motifs < 1024 ~ "Leaf 3",
                          transitivity >= 0.0207311636201516 & motifs >= 1024 ~ "Leaf 4"),
         leaf = factor(leaf))
```

@@

Summary of amenity categories by DA and Type:

```{r}
amenities_walksheds_summary <- amenities_walksheds |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

To map accessibility by category of amenity we first categorize the number of amenities:

```{r}
amentities_da <-  urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE))
```

Plot:

```{r}
amentities_da |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Pivot wider:

```{r}
junk <- amentities_da |>
  pivot_wider(names_from = "Amenity_Category",
              values_from = "Number_of_Amenities")
```
